---
title: "Problem Set 1"
author: "Edgar Aguilar"
date: January 29
output: pdf_document
---

```{r setup, include=FALSE}
# This "setup" chunk specifies global options
# for handling code, plots, etc in your doc.
  knitr::opts_chunk$set(
    eval = TRUE,
    echo = TRUE, 
    warning = FALSE, 
    message = FALSE,
    fig.align = 'center'
  )  

# This is the appropriate place to load packages,
# import data, set themes, and more
  library(tidyverse)
  library(kableExtra)

# Import the Afrobarometer dataset
df1 <- read_csv('output file.csv')

```


# Answers to questions for PS1

## Question #1
- I am using the data set from Congo-Brazzaville. 
- The survey includes information from `r length(unique(df1$RESPNO))` respondents.
- The responses were collected between `r min(df1$DATEINTR)` and `r max(df1$DATEINTR)`.

## Question #2
The ages were between `r min(df1$Q1)` and `r max(df1$Q1)`, with a mean of `r mean(df1$Q1)` and median of `r median(df1$Q1)`. The other characteristics of the respodants are in the table below: 

``` {r, echo = TRUE, fig.cap='Descriptive Statistics Summary'}
# Use kableExtra to show a final nicer table
library(kableExtra)

## Gender ----
# Show unique values in the GENDER column -- 1 = Man 2 = Woman. -- to check data looks good.

# Replace 1 with "Man" and 2 with "Woman", and calculate percentages. Generate tible.
gender_tibble <- df1 %>%
  mutate(
    Q100 = case_when(
      Q100 == 1 ~ "Man",
      Q100 == 2 ~ "Woman",
    )
  ) %>%
  count(Q100, name = "Count") %>%  
  mutate(Percentage = round((Count / sum(Count)) * 100, 2), Category = "Gender")
##category variable will help me put tibles together in a single table after.


## Region ----
## replace numbers with region names bases on codebook and create tible
region_tibble <- df1 %>%
  mutate(
    REGION = case_when(
      REGION == 1780 ~ "Brazzaville",
      REGION == 1781 ~ "Pointe-Noire",
      REGION == 1782 ~ "Kouilou",
      REGION == 1783 ~ "Niari",
      REGION == 1784 ~ "Bouenza",
      REGION == 1785 ~ "Lekoumou",
      REGION == 1786 ~ "Pool",
      REGION == 1787 ~ "Plateaux",
      REGION == 1788 ~ "Cuvette",
      REGION == 1789 ~ "Cuvette Ouest",
      REGION == 1790 ~ "Sangha",
      REGION == 1791 ~ "Likouala ",
    )
  ) %>%
  count(REGION, name = "Count") %>%
  mutate(Percentage = round((Count/sum(Count))*100, 2), Category = "Region")


## Location ----
# Replace 1 with "Urban" and 2 with "Rural", and calculate percentages
urbrur_tibble <- df1 %>%
  mutate(
    URBRUR = case_when(
      URBRUR == 1 ~ "Urban",
      URBRUR == 2 ~ "Rural",
    )
  ) %>%
  count(URBRUR, name = "Count") %>%  
  mutate(Percentage = round((Count / sum(Count)) * 100, 2), Category = "Location") 


## Language ----
# Replace and calculate 
language_tibble <- df1 %>%
  mutate(
    Q2 = case_when(
      Q2 == 2 ~ "Français",
      Q2 == 1780 ~ "Lingala",
      Q2 == 1781 ~ "Kituba",
      Q2 == 1782 ~ "Lari",
      Q2 == 1783 ~ "Teke",
      Q2 == 9995 ~ "Other",
      Q2 == 9998 ~ "A refusé de répondre",
      Q2 == 9999 ~ "Ne sait pas",
      Q2 == -1 ~ "Manquant ",
    )
  ) %>%
  count(Q2, name = "Count") %>%  
  mutate(Percentage = round((Count / sum(Count)) * 100, 2), Category = "Language") 


## Summary Disriptive Statistics ----
# combine tibbles
summary_table <- bind_rows(
  gender_tibble,
  region_tibble,
  urbrur_tibble,
  language_tibble
)

# Merge Q100, REGION, URBRUR, and Q2 into one column called 'Variable'
updated_summary_table <- summary_table %>%
  mutate(
    Variable = coalesce(as.character(Q100), as.character(REGION), as.character(URBRUR), as.character(Q2))  # Combine columns into 'Variable'
  ) %>%
  select(Category, Variable, Count, Percentage)  # Rearrange columns to the desired order

# Display the updated table
updated_summary_table %>%
  knitr::kable(caption = "Descriptive Statistics Summary Table") %>%
  kable_styling(full_width = FALSE)
```

## Question #3
- 21% of respondents perceive China’s economic and political influence as negative while 68% see it as positive.
- 9% consider China's influence neither positive nor negative and around 2% are unsure.
```{r, echo = TRUE, fig.cap='Attitudes toward China'}
# create frequency table
China_attitudes_frequency <- df1 %>%
  count(Q78A, name = "Count") %>%
  mutate(
    Percentage = round((Count / sum(Count)) * 100, 2),
    Response = case_when(
      Q78A == 1 ~ "Very negative",
      Q78A == 2 ~ "Somewhat negative",
      Q78A == 3 ~ "Neither positive nor negative",
      Q78A == 4 ~ "Somewhat positive",
      Q78A == 5 ~ "Very positive",
      Q78A == 8 ~ "Refused",
      Q78A == 9 ~ "Don't know",
      Q78A == -1 ~ "Missing",
    )
  ) %>%
  select(Response, Count, Percentage)

# Display the table China
China_attitudes_frequency %>%
  knitr::kable(
    caption = "Attitudes toward China Table" 
  ) %>%
  kable_styling(
    full_width = FALSE
  )
```

## Question #4
- 21% of respondents perceive USA’s economic and political influence as negative while 51% see it as positive.
- 25% of respondents consider it neither positive nor negative and 2% do not know. 
- The perception of US influence as positive is 17% lower than China, but there is a greater percentage of people (16% more in comparison with China) that consider it neither positive or negative. 
```{r, echo = TRUE, fig.cap='Attitudes toward USA'}
#create frequency table
USA_attitudes_frequency <- df1 %>%
  count(Q78B, name = "Count") %>%
  mutate(
    Percentage = round((Count / sum(Count)) * 100, 2),
    Response = case_when(
      Q78B == 1 ~ "Very negative",
      Q78B == 2 ~ "Somewhat negative",
      Q78B == 3 ~ "Neither positive nor negative",
      Q78B == 4 ~ "Somewhat positive",
      Q78B == 5 ~ "Very positive",
      Q78B == 8 ~ "Refused",
      Q78B == 9 ~ "Don't know",
      Q78B == -1 ~ "Missing",
    )
  ) %>%
  select(Response, Count, Percentage)

# Display the table USA
USA_attitudes_frequency %>%
  knitr::kable(
    caption = "Attitudes Toward USA Table" 
  ) %>%
  kable_styling(
    full_width = FALSE
  )
```

## Question #5
- The results show that, on average, responses for the USA are 0.304 points higher than those for China. So, on average respondents perceive the US influence in economics and politics more favorably than China's. The high t-value (8.8) indicates that there is a significant difference between responses to China and the USA and the low p-value (below 0) suggests that the results are statistically significant. We reject null hypothesis that there is not difference in the means. 
```{r, echo = TRUE, fig.cap='T-test'}
# Clean Q78A and Q78B using across
df_clean <- df1 %>%
  mutate(
    across(
      Q78A:Q78B,  
      ~ if_else(.x %in% 1:5, .x, NA)
    )
  )

# Perform the paired t-test
t_test_result <- t.test(df_clean$Q78A, df_clean$Q78B, paired = TRUE)

# Display the t-test result
t_test_result 
```
