---
title: "Problem Set 1"
author: "Edgar Aguilar"
date: January 24
output: pdf_document
---

```{r setup, include=FALSE}
# This "setup" chunk specifies global options
# for handling code, plots, etc in your doc.
  knitr::opts_chunk$set(
    eval = TRUE,
    echo = TRUE, 
    warning = FALSE, 
    message = FALSE,
    fig.align = 'center'
  )  

# This is the appropriate place to load packages,
# import data, set themes, and more
  library(tidyverse)
  library(kableExtra)

# Import the Afrobarometer dataset
df1 <- read_csv('output file.csv')

```


# Answers to questions

## Question #1
- I am using the data set from Congo-Brazzaville. 
- The survey includes information from `r length(unique(df1$RESPNO))` respondents.
- The responses were collected between `r min(df1$DATEINTR)` and `r max(df1$DATEINTR)`.

## Question #2
The ages were between `r min(df1$Q1)` and `r max(df1$Q1)`, with a mean of `r mean(df1$Q1)` and median of `r mean(df1$Q1)`. The other characteristics of the respodants are in the table below. 

``` {r, echo = FALSE, fig.cap='Descriptive Statistics Summary'}
# Display the updated summary table without showing the code
library(kableExtra)

## Gender ----
# Show unique values in the GENDER column -- 1 = Man 2 = Woman. -- to check data looks good.

# Replace 1 with "Man" and 2 with "Woman", and calculate percentages
gender_tibble <- df1 %>%
  mutate(
    Q100 = case_when(
      Q100 == 1 ~ "Man",
      Q100 == 2 ~ "Woman",
    )
  ) %>%
  count(Q100, name = "Count") %>%  
  mutate(Percentage = round((Count / sum(Count)) * 100, 2), Category = "Gender")


## Region ----


region_tibble <- df1 %>%
  mutate(
    REGION = case_when(
      REGION == 1780 ~ "Brazzaville",
      REGION == 1781 ~ "Pointe-Noire",
      REGION == 1782 ~ "Kouilou",
      REGION == 1783 ~ "Niari",
      REGION == 1784 ~ "Bouenza",
      REGION == 1785 ~ "Lekoumou",
      REGION == 1786 ~ "Pool",
      REGION == 1787 ~ "Plateaux",
      REGION == 1788 ~ "Cuvette",
      REGION == 1789 ~ "Cuvette Ouest",
      REGION == 1790 ~ "Sangha",
      REGION == 1791 ~ "Likouala ",
    )
  ) %>%
  count(REGION, name = "Count") %>%
  mutate(Percentage = round((Count/sum(Count))*100, 2), Category = "Region")


## Location ----

# Replace 1 with "Urban" and 2 with "Rural", and calculate percentages
urbrur_tibble <- df1 %>%
  mutate(
    URBRUR = case_when(
      URBRUR == 1 ~ "Urban",
      URBRUR == 2 ~ "Rural",
    )
  ) %>%
  count(URBRUR, name = "Count") %>%  
  mutate(Percentage = round((Count / sum(Count)) * 100, 2), Category = "Location") 


## Language ----


# Replace and calculate categories
language_tibble <- df1 %>%
  mutate(
    Q2 = case_when(
      Q2 == 2 ~ "Français",
      Q2 == 1780 ~ "Lingala",
      Q2 == 1781 ~ "Kituba",
      Q2 == 1782 ~ "Lari",
      Q2 == 1783 ~ "Teke",
      Q2 == 9995 ~ "Other",
      Q2 == 9998 ~ "A refusé de répondre",
      Q2 == 9999 ~ "Ne sait pas",
      Q2 == -1 ~ "Manquant ",
    )
  ) %>%
  count(Q2, name = "Count") %>%  
  mutate(Percentage = round((Count / sum(Count)) * 100, 2), Category = "Language") 


## Summary Disriptive Statistics ----
# combine tibbles
summary_table <- bind_rows(
  gender_tibble,
  region_tibble,
  urbrur_tibble,
  language_tibble
)

# Merge Q100, REGION, URBRUR, and Q2 into one column called 'Variable'
updated_summary_table <- summary_table %>%
  mutate(
    Variable = coalesce(as.character(Q100), as.character(REGION), as.character(URBRUR), as.character(Q2))  # Combine columns into 'Variable'
  ) %>%
  select(Category, Variable, Count, Percentage)  # Rearrange columns to the desired order

# Display the updated table
updated_summary_table %>%
  knitr::kable(caption = "Descriptive Statistics Summary Table") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

## Question #3

```{r, echo = FALSE, fig.cap='Attitudes toward China'}
China_attitudes_frequency <- df1 %>%
  count(Q78A, name = "Count") %>%
  mutate(
    Percentage = round((Count / sum(Count)) * 100, 2),
    Response = case_when(
      Q78A == 1 ~ "Very negative",
      Q78A == 2 ~ "Somewhat negative",
      Q78A == 3 ~ "Neither positive nor negative",
      Q78A == 4 ~ "Somewhat positive",
      Q78A == 5 ~ "Very positive",
      Q78A == 8 ~ "Refused",
      Q78A == 9 ~ "Don't know",
      Q78A == -1 ~ "Missing",
    )
  ) %>%
  select(Response, Count, Percentage)

# Display the table China
China_attitudes_frequency %>%
  knitr::kable(
    caption = "China Attitudes Frequency Table"  # Add a meaningful caption
  ) %>%
  kable_styling(
    full_width = FALSE, 
    bootstrap_options = c("striped", "hover")  # Add styling options
  )
```

## Question #4

```{r, echo = FALSE, fig.cap='Attitudes toward USA'}
USA_attitudes_frequency <- df1 %>%
  count(Q78B, name = "Count") %>%
  mutate(
    Percentage = round((Count / sum(Count)) * 100, 2),
    Response = case_when(
      Q78B == 1 ~ "Very negative",
      Q78B == 2 ~ "Somewhat negative",
      Q78B == 3 ~ "Neither positive nor negative",
      Q78B == 4 ~ "Somewhat positive",
      Q78B == 5 ~ "Very positive",
      Q78B == 8 ~ "Refused",
      Q78B == 9 ~ "Don't know",
      Q78B == -1 ~ "Missing",
    )
  ) %>%
  select(Response, Count, Percentage)

# Display the table China
USA_attitudes_frequency %>%
  knitr::kable(
    caption = "USA Attitudes Frequency Table" 
  ) %>%
  kable_styling(
    full_width = FALSE, 
    bootstrap_options = c("striped", "hover")
  )
```

## Question #5

```{r, echo = FALSE, fig.cap='T-test'}
# Clean Q78A and Q78B using across
df_clean <- df1 %>%
  mutate(
    across(
      Q78A:Q78B,  
      ~ if_else(.x %in% 1:5, .x, NA)
    )
  )

# Perform the paired t-test
t_test_result <- t.test(df_clean$Q78A, df_clean$Q78B, paired = TRUE)

# Clean Q78A and Q78B using across
df_clean <- df1 %>%
  mutate(
    across(
      Q78A:Q78B,  
      ~ if_else(.x %in% 1:5, .x, NA)
    )
  )

# Perform the paired t-test
t_test_result <- t.test(df_clean$Q78A, df_clean$Q78B, paired = TRUE)

# Display the t-test result
t_test_result 
```
